---
title: 'Vignette for eigTest'
author: "Yuchen Xu"
date: "11/10/2020"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette for eigTest}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package installation

To install and load the package, use the code:
```{r install, message=FALSE, warning=FALSE}
devtools::install_github('XycYuchenXu/eigTest', quiet = T)
library(eigTest)
```

## Simulation samples

The sample generating process is as follows:

1. Randomly sample a common eigenvector matrix $V \in \mathbb{R}^{d \times d}$ where $d$ is the dimension of the square matrix and the entries are i.i.d. standard normal;

2. Adding noise to the matrix $V$ to get $V_i(\sigma) = V + \sigma E_i$ for $i = 1, \dots, p$, where signal to noise ratio (SNR) is defined as $\frac{1}{\sigma^2}$) and entries in $E_i$ are i.i.d. standard normal.

3. Randomly sample vectors $d_i \in \mathbb{R}^d$ with uniform i.i.d. entries in interval $(-1,1)$, and generate square matrices $\mu_i(\sigma) = V_i^{-1}(\sigma) D_i V_i(\sigma)$ where $D_i = \mbox{diag}(d_i)$ is a diagonal matrix.

4. Given sample size $n$, generate $n$ random matrices $X_{ij}, j = 1, \dots, n$ with mean $\mu_i(\sigma)$ and identity variance matrix, and get the estimates $A_i = \bar X_{i}$, $\Sigma_i = \frac{1}{n-1} \sum_{j=1}^n \mbox{vec}(X_{ij} - A_i) \mbox{vec}'(X_{ij} - A_i)$.

The package includes functions to generate simulation samples in the following way:

```{r simuSamples}
twoSample = generateMeans(d = 5,           # square matrix's dimension
                          p = 2,           # number of matrices
                          snr = c(10,1))   # vector of SNR in eigenvector
```


## Functionality

### Two-sample test

```{r twoSample, eval=FALSE, include=FALSE}
## Commutator-based test
commutatorTest()

## Log-likelihood ratio test
projTest()
```

### Common eigenvector estimate

```{r eigEst, eval=FALSE, include=FALSE}
JDTE()
```

### Multi-sample test

```{r multiSample, eval=FALSE, include=FALSE}
eigTest()
```

### Partial test

```{r partial, eval=FALSE, include=FALSE}

```

